### 2023-11-30 main.tex
S2

@@

\usepackage[margin=2.5cm]{geometry}

@@

In closing this introduction we mention that we are currently working on the boundary-driven multispecies 
model with non-compact Lie algebras, both integrable and non-integrable, which will be presented in a 
future paper.

@@

\noindent
\begin{remark}
\label{hole}
In the following we will often consider type $N$ as a distinguished type
(called `holes'), whose number at each site is fixed by the numbers of particles 
of the remaining $N-1$ types. It is interesting to observe that, in case we stop distinguishing the types of the ``true'' particles of types 
(those of types $\{1,\ldots,N-1\}$), we retrieve the 
boundary-driven version of the partial symmetric exclusion process \cite{schutzSandow,carinci2013duality}.
The choice of type $N$ for the distinguished type is arbitrary and does not affect
the results. In particular, we shall see in Section~\ref{sectionDuality} about duality, that
the dual process eventually has only particles of type $N$ on the graph.
Interpreting type $N$ as an hole, one can then say that the dual process voids the graph.  
\end{remark}

@@

with marginals $\mu_{\text{rev}}^{x}$ given by 

@@

each reservoir at site $x$ wants to fix its own set of densities $\rho^{x}=(\rho_{1}^{x},\ldots,\rho_{N}^{x})$ with

@@

@@

In this paper we will often use the fact that the Markov generator of the multispecies stirring process can be described in terms a Lie algebra of rank $N$.
In this section we provide details about this.

Consider the Lie algebra ${gl}(N)$ with generators denoted by $\EE_{ab}$ with $a,b\in \{1,\ldots,N\}$ and commutation relations
\begin{equation}\label{eq:comgl}
\left[\EE_{ab},\EE_{cd}\right]=\EE_{ad}\delta_{cb}-\EE_{cb}\delta_{ad}\qquad \forall a,b\in \{1,\ldots,N\}\,.
\end{equation}
The finite-dimensional representations are labelled by partitions $\lambda=(\lambda_1,\lambda_2,\ldots,\lambda_N)$ of $\nu$ with  $\lambda_i\geq \lambda_{i+1}$,  $\lambda_i\in \mathbb{N}$ and $\sum_{i=1}^N \lambda_i = \nu\in\mathbb{N}$.  
We are interested in the {\em symmetric} finite-dimensional representations with 
\begin{equation}\label{eq:dynkin}
    \lambda=(\twoj,0,\ldots,0) \;.
    %\qquad\text{where}\qquad \twoj\in\mathbb{N}\,.
\end{equation} 
The dimension $M_\twoj$ of this symmetric representations is given by the combination of $N$ objects in $\twoj$ positions with repetition, namely
\begin{equation}
	M_\twoj= \frac{(N+\twoj-1)!}{\twoj  !(N-1)!}\dt
\end{equation} 
The generators of the symmetric representations will be denoted by $E_{ab}$.
A basis of the vector space $\mathbb{C}^{M_\twoj}$ are the column vectors denoted by
\begin{equation}
  |n\rangle=  |n_{1},\ldots,n_{N}\rangle,\quad \text{with}\quad n_{i}\in\mathbb{N}_{0}\quad \text{sucht that}\quad \sum_{i=1}^{N}n_{i}=\nu\,.
\end{equation}
%where $\sum_{a=1}^{N}n_{i}=\nu$. 
The basis vectors satisfy the orthogonality relation
\begin{equation}\label{ortho}
   \langle m|n \rangle =\langle m_{1},\ldots,m_{N}|n_{1},\ldots,n_{N}\rangle=\prod_{a=1}^{N}\delta_{m_{a},n_{a}}\co
\end{equation}
where  $ \langle m_{1},\ldots,m_{N}|$ is the row vectors obtained by transposing $|m\rangle=|m_{1},\ldots,m_{N}\rangle$ and $\delta_{m_{a},n_{a}}$ is the Kronecker delta.

@@

The matrices defined in this way satisfy the commutation relations \eqref{eq:comgl} and yield Dynkin weight \eqref{eq:dynkin}. \\

@@

The process with generator \eqref{Generator} can be described in terms of ${gl}(N)$ Lie algebra generators. The state space \eqref{stateSpace} is given by the $|V|$-fold tensor product of the vector space with basis elements $|n^x\rangle$ at a given site. Namely,

@@

with $\sum_{a=1}^{N}n_{a}^{x}=\nu$ for any $x\in V$. Sometimes it will bw convenient to write $|n^{x}\rangle$ to denote, for a fixed $x\in V$, the vector $|n_{1}^{x},\ldots,n_{N}^{x}\rangle$.
The following orthogonality relation is a consequence of the single site relation \eqref{ortho}

@@

where the edge Hamiltonian that describes the interaction between two connected sites is

@@

  and where the site Hamiltonian is

@@

Here $E_{ab}^{x}$ is a copy of $E_{ab}$ defined in \eqref{actionE} acting on site $x$ (and
acting trivially as the identity on the other sites). 
Following  \cite{belitsky2015self}, the Hamiltonian and the Markov generator are linked by

@@

H=\mathcal{L}^{T}\dt
The action of the generator can also be expressed as 

@@

that acts diagonally as $C|\bm{n}\rangle=\twoj(\twoj+N)|\bm{n}\rangle$ and belogs to the center of ${gl}(N)$ (i.e. it commutes with all the algebra elements).  

@@

\Delta:{gl}(N)&\to {gl}(N)\otimes {gl}(N)\\
E_{ab}&\mapsto E_{ab}\otimes \mathbbm{1}+\mathbbm{1}\otimes E_{ab}

>>>>

% \usepackage[margin=2.5cm]{geometry}

%%%%%  TODO NOTES ETC  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\usepackage[left=1.7cm,right=4.3cm,bottom=4cm]{geometry}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\definecolor{todocolor}{HTML}{D7E1E5}
\usepackage[textsize=tiny, 
            backgroundcolor=todocolor, 
            bordercolor=todocolor, 
            linecolor=todocolor, 
            textwidth=3.9cm]{todonotes}
\newcommand{\todoWARN}[1]{\todo{\colorbox{yellow}{\parbox{3.4cm}{#1}}}}
\newcommand{\todoERR}[1]{\todo{\colorbox{red}{\parbox{3.4cm}{#1}}}}
\usepackage{comment}
\usepackage{rotating}
\usepackage[left]{showlabels}
\renewcommand{\showlabelfont}{\tiny}
\renewcommand{\showlabelsetlabel}[1]
{\begin{turn}{65}\showlabelfont \llap{#1}\end{turn}}
\newcommand{\wiptitle}[1]{
    \par\noindent\hspace*{-0.5cm}\texttt{\scriptsize\color{gray} \lbrack #1\rbrack} \par\noindent\ignorespaces}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

@@

In closing this introduction we mention that we are currently working on the boundary-driven multispecies 
model with non-compact Lie algebras, both integrable and non-integrable, which will be presented in a 
future paper. 
\todo{I prefer to remove this sentence}

@@

\noindent
\begin{remark}
\label{hole}
In the following we will often consider type $N$ as a distinguished type
(called `holes'), whose number at each site is fixed by the numbers of particles 
of the remaining $N-1$ types. 
% It is interesting to observe that, in case we stop distinguishing the types of the ``true'' particles of types 
% (those of types $\{1,\ldots,N-1\}$), we retrieve the 
% boundary-driven version of the partial symmetric exclusion process \cite{schutzSandow,carinci2013duality}.
In the case $N=2$, we retrieve the  boundary-driven version of the symmetric exclusion process \cite{schutzSandow,carinci2013duality}.
% \todo{I would prefer to say this is for N=2}
The choice of type $N$ for the distinguished type is arbitrary and does not affect
the results. However, we shall see in Section~\ref{sectionDuality}, that
the dual process breaks the symmetry.
In particular, in the  long-time limit the dual process has only particles of type $N$ on the graph.
Interpreting type $N$ as an hole, one can then say that the dual process voids the graph.  
\todo{Exchange hole with void? Also can the remark be just text?}
\end{remark}

@@

with marginals $\mu_{\text{rev}}^{x}$ given by  the multinomial distribution

@@

each reservoir at site $x$ {\color{red}wants to fix}\todo{?} its own set of densities $\rho^{x}=(\rho_{1}^{x},\ldots,\rho_{N}^{x})$ with

@@

		\todo{?}

@@

In this paper we will often use the fact that the Markov generator of the multispecies stirring process can be described in terms a Lie algebra ${gl}(N)$.
In this section we provide details about this.

Consider the Lie algebra ${gl}(N)$ with generators denoted by $\EE_{ab}$ with $a,b\in \{1,\ldots,N\}$ and commutation relations
\begin{equation}\label{eq:comgl}
\left[\EE_{ab},\EE_{cd}\right]=\EE_{ad}\delta_{cb}-\EE_{cb}\delta_{ad}\qquad \forall a,b\in \{1,\ldots,N\}\,.
\end{equation}
The finite-dimensional representations are labelled by partitions $\lambda=(\lambda_1,\lambda_2,\ldots,\lambda_N)$ of $\nu$ with  $\lambda_i\geq \lambda_{i+1}$,  $\lambda_i\in \mathbb{N}$ and $\sum_{i=1}^N \lambda_i = \nu\in\mathbb{N}$.  
We are interested in the {\em symmetric} finite-dimensional representations with 
\begin{equation}\label{eq:dynkin}
    \lambda=(\twoj,0,\ldots,0) \;.
    %\qquad\text{where}\qquad \twoj\in\mathbb{N}\,.
\end{equation} 
The dimension $M_\twoj$ of this symmetric representations is given by the combination of $N$ objects in $\twoj$ positions with repetition, namely
\begin{equation}
	M_\twoj= \frac{(N+\twoj-1)!}{\twoj  !(N-1)!}\dt
\end{equation} 
The generators of the symmetric representations will be denoted by $E_{ab}$.
A basis of the vector space $\mathbb{C}^{M_\twoj}$ are the column vectors denoted by
\begin{equation}
  |n\rangle=  |n_{1},\ldots,n_{N}\rangle,\quad \text{with}\quad n_{i}\in\mathbb{N}_{0}\quad \text{such that}\quad \sum_{i=1}^{N}n_{i}=\nu\,.
\end{equation}
%where $\sum_{a=1}^{N}n_{i}=\nu$. 
The basis vectors satisfy the orthogonality relation
\begin{equation}\label{ortho}
   \langle m|n \rangle =\langle m_{1},\ldots,m_{N}|n_{1},\ldots,n_{N}\rangle=\prod_{a=1}^{N}\delta_{m_{a},n_{a}}\co
\end{equation}
where  $ \langle m_{1},\ldots,m_{N}|$ is the row vector obtained by transposing $|m\rangle=|m_{1},\ldots,m_{N}\rangle$ and $\delta_{m_{a},n_{a}}$ is the Kronecker delta.

@@

The matrices defined in this way satisfy the commutation relations \eqref{eq:comgl} and yield highest Dynkin weight \eqref{eq:dynkin}. \\

@@

As mentioned above, the process with generator \eqref{Generator} can be described in terms of ${gl}(N)$ Lie algebra generators. The state space \eqref{stateSpace} is given by the $|V|$-fold tensor product of the vector space with basis elements $|n^x\rangle$ at a given site. Namely,

@@

with $\sum_{a=1}^{N}n_{a}^{x}=\nu$ for any $x\in V$. For a fixed $x\in V$ we write $|n^{x}\rangle=|n_{1}^{x},\ldots,n_{N}^{x}\rangle$. The following orthogonality relation is a consequence of the single site relation \eqref{ortho}

@@

where the edge Hamiltonian $\mathcal{H}_{x,y}$ that describes the interaction between two connected sites is

@@

  and where the site Hamiltonian $H_{x}$  is

@@

Here $E_{ab}^{x}$ denotes the generator $E_{ab}$ in \eqref{actionE} acting non-trivially on site $x$ (and as the identity on the other sites). 
The Hamiltonian in \eqref{OriginalHamiltonian} is stochastic and is linked to the Markov generator  by

@@

H=\mathcal{L}^{T} 
where $T$ denotes transposition, see e.g.  \cite{belitsky2015self}.
The action of the generator on functions $f$ can then  be expressed as 

@@

that acts diagonally as $C|{n}\rangle=\twoj(\twoj+N)|{n}\rangle$ and belogs to the center of ${gl}(N)$ (i.e. it commutes with all the algebra elements).  

@@

\Delta:{gl}(N)\to {gl}(N)\otimes {gl}(N)
E_{ab}\mapsto E_{ab}\otimes \mathbbm{1}+\mathbbm{1}\otimes E_{ab}\co
